Refinement of Parallel Algorithms down to LLVM

### 2024-03-25 papers/JAR_SI_ITP2022/main.tex
last changes before revision submission

The formalization is available at \href{https://www21.in.tum.de/~lammich/isabelle_llvm_par/}{www21.in.tum.de/$\sim$lammich/isabelle\_llvm\_par}.

@@

    interpretation idx: pcmp lt_idx lt_idx$_\dagger$ arr$_A$ idx$_A$   \<proof>
  \end{lstlisting}
  This yields sorting algorithms for sorting indexes, taking an extra parameter for the array to index into.
  For our sampling application, we use \is{idx.introsort xs}.

@@

    \subsection{Statistics}
    The Isabelle Refinement Framework for LLVM consists of roughly 45k lines of theory text and Isabelle-ML code (referred to as kLOC from now on). The sorting algorithms comprise another 14kLOC.

@@

      An important aspect is the scalability of our tools. The current implementation scales to small software, like the verified IsaSAT-solver~\cite{FlLa23}, which actually uses our sequential sorting algorithms. However, for projects of this size, the times needed by Sepref and the LLVM code exporter easily get into the order of dozens of minutes, and some manual performance tweaks are required (cf.~\cite{FlLa23}[Sec.~5]).
      Further improvement of the scalability is left to future work.

>>>>

Isabelle LLVM is hosted at \href{https://lammich.github.io/isabelle_llvm/index.html}{lammich.github.io/isabelle\_llvm}.
The version described in this paper has been archived~\cite{Isabelle_LLVM_artefact}.
%
% The formalization is available at \href{https://www21.in.tum.de/~lammich/isabelle_llvm_par/}{www21.in.tum.de/$\sim$lammich/isabelle\_llvm\_par}.

@@

    interpretation idx: pcmp lt_idx lt_idx$_\dagger$ arr$_A$ idx$_A$   \<proof>
  \end{lstlisting}
  This instantiates the generic sorting algorithms defined in the \is{pcmp} locale to use \is{lt_idx xs} as comparison function, taking \is{xs} as an extra parameter. To sort our list \is{$is$} of sample indexes into \is{xs}, we use the instantiation of the introsort algorithm: \is{idx.introsort xs $is$}.

@@

    \subsection{Development Effort}
    The Isabelle Refinement Framework for LLVM consists of roughly 45k lines of theory text and Isabelle-ML code (referred to as kLOC from now on). The sorting algorithms comprise another 14kLOC.

@@

      An important aspect is the scalability of our tools. The current implementation scales to small software systems, like the verified IsaSAT-solver~\cite{FlLa23}, which actually uses our sequential sorting algorithms. However, for projects of this size, the times needed by Sepref and the LLVM code exporter easily get into the order of dozens of minutes, and some manual performance tweaks are required (cf.~\cite{FlLa23}[Sec.~5]).
      Further improvement of the scalability is left to future work.
